---
title: "Analyzing Clean Eating Month Results"
output:
  flexdashboard::flex_dashboard: null
  orientation: columns
  vertical_layout: fill
social: menu
source_code: embed
runtime: shiny
resource_files:
- raw_data/data.csv
- raw_data/data.csv
---
  
```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(glue)
library(shiny)

# global constants
RAW_DATA_DIR <- "./raw_data/"
RAW_DATA_FILE_NAME <- "data.csv"

# read the data
df <- read_csv(file.path(RAW_DATA_DIR, RAW_DATA_FILE_NAME))

# remove some fields for which we know we dont have data
df <- df %>%
  select(-`muscle-percentage`, -`lean-mass`) %>%
  drop_na() %>%
  arrange(date)

# determine the per day weight loss dataframe by
# calculating loss as weight - the one previous value of weight
# this is done by first grouping the dataframe by name since it has
# data for two people and then arranging by date while maintaining
# the grouping (not the .by_group=TRUE)
df_wt_loss <- df %>%
  select(name, date, weight) %>%
  group_by(name) %>%
  arrange(date, .by_group=TRUE) %>%
  mutate(loss_per_day = weight-lag(weight, 1)) 

# is the curse of the weekend real? Assign the day to each date
df_wt_loss <- df_wt_loss %>%
  mutate(day = weekdays(as.Date(date)))
```

Data
=====================================  

Column {data-width=500}
-----------------------------------------------------------------------
  
### How did we do? Just give me the easy plots.
```{r}
renderPlot({
  # convert  the data to tidy (long) format
  # also rename some columns to add unit
  df_tidy <- df %>%
    rename(`bmi (kg/m^2)`=bmi, `weight (pounds)`=weight) %>%
    gather(metric, value, -date, -name)
  
  # we would like to arrange the charts in an order
  # such that weight and bmi appear side by side
  # easiest way to do this is to convert the name
  # to a factor and then manually arrange in the desired
  # order
  df_tidy$metric <- factor(df_tidy$metric,
                           levels = c("weight (pounds)",
                                      "bmi (kg/m^2)",
                                      "total-lean-mass-percentage",
                                      "body-fat-percentage",
                                      "hydration-percentage"))
  df_tidy <- df_tidy %>%
    arrange(metric)
  
  # plot this as a line plot + smoothing line using linear model
  # facet it with the metric name, color it by person name
  df_tidy %>%
    ggplot(aes(x=date, y=value, col=name)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~metric, scales="free_y") + 
    xlab("") +
    ylab("") +
    theme_bw() +
    theme(legend.position = "bottom",
          legend.title = element_blank(),
          strip.text.x = element_text(size = 12)) +
    scale_color_tableau()
})

```


Column {data-width=500}
-----------------------------------------------------------------------
### I need to see the raw data, show me the numbers not the interpretation!

```{r}
renderTable({
  # display the raw data as a table, rename column names
  # to friendly name
  df %>%
    select(-`hydration-percentage`) %>%
    arrange(desc(date)) %>%
    mutate(date = as.character(date)) %>%
    rename(Name=name, Date=date, `Weight (pounds)`=weight,
           `BMI (kg/m^2)`=bmi,
           `Body Fat (%)`=`body-fat-percentage`,
           `Lean mass (%)`=`total-lean-mass-percentage`)
}, striped=TRUE, hover=TRUE)

```


Hypothesis {data-orientation=rows}
=====================================     
   
Row {data-height=500}
-------------------------------------

### Is the rate of weight loss slowing down?
Seems like weight loss fatigue is setting as the body had adjusted to the diet and exercise regime.
```{r}
renderPlot({
  # plot a timeseries of the per day weight loss
  # alongwith a smoothing line to show the trend
  df_wt_loss %>%
    ggplot(aes(x=date, y=loss_per_day, fill=name, color=name)) +
    geom_point() + 
    geom_smooth(method = "lm") +
    theme_bw() +
    xlab("") +
    ylab("weight loss in pounds (negative represents loss)") +
    theme(legend.position = "right",
          legend.title = element_blank(),
          strip.text.x = element_text(size = 12)) + 
    scale_color_tableau() + 
    scale_fill_tableau()
})
```

### Is there such a thing as the "curse of the weekend"?
Looks like it is true, there is a weight gain on Sunday and only a marginal loss on Monday.
```{r}
renderPlot({
  # plot a per day side by side (for each person)
  # bar chart showing the median value of the weight loss
  # on each day of the week
  df_wt_loss_by_day <- df_wt_loss %>%
    group_by(name, day) %>%
    summarize(median_wt_loss = median(loss_per_day, na.rm=T))
  
  # arrange the week days in order
  df_wt_loss_by_day$day <- factor(df_wt_loss_by_day$day,
                                  levels = c("Monday",
                                             "Tuesday",
                                             "Wednesday",
                                             "Thursday",
                                             "Friday",
                                             "Saturday",
                                             "Sunday"))
  
  # plot the bar plot with each day having two side by side bars
  # not use of stat="identity" since we have the y value for each
  # bar and position=dodge to show side by side bars
  df_wt_loss_by_day %>%
    arrange(day) %>%
    ggplot(aes(x=day, y=median_wt_loss, fill=name)) +
    geom_bar(stat="identity", position="dodge")  + 
    xlab("") + 
    ylab("Median weight loss (pounds)") + 
    scale_color_tableau() +
    scale_fill_tableau() +
    theme_bw() + 
    theme(legend.title = element_blank())
})
```

Row {data-height=500}
-------------------------------------
   
### Is there a difference in the weight loss patterns for Nidhi and Amit
Amit's per day weight loss has a lot of variation as compared to Nidhi's.
```{r}
renderPlot({
  # median value of per day weight loss
  median_wt_loss_per_day_amit <- round(median(df_wt_loss[df_wt_loss$name == "Amit", ]$loss_per_day, na.rm=TRUE), 2)
  
  median_wt_loss_per_day_nidhi <- round(median(df_wt_loss[df_wt_loss$name == "Nidhi", ]$loss_per_day, na.rm=TRUE), 2)

  # density plot, col=name so that we get two density plots, one for each person
  p <- df_wt_loss %>%
    drop_na() %>%
    ggplot(aes(loss_per_day, col=name, fill=name)) + 
    geom_density(alpha = 0.1) +
    xlab("") + 
    ylab("Density") + 
    scale_color_tableau() +
    scale_fill_tableau() +
    theme_bw() + 
    theme(legend.position = "right", legend.title = element_blank()) 
  
  # add a text label to the plot at a convenient location (trial and error)
  # to show median value
  p + geom_label(
      label=glue("Median weight loss per day:\n Amit={median_wt_loss_per_day_amit} pounds,   Nidhi={median_wt_loss_per_day_nidhi} pounds"), 
      x=-2.5,
      y=0.35,
      label.padding = unit(0.55, "lines"), # Rectangle size around label
      label.size = 0.35,
      color = "black",
      fill="#69b3a2")
})
```   
    
### How do Nidhi and Amit compare on every metric?
Amit lost more weight in absolute terms but on other metrics Nidhi performed better.
```{r}
renderPlot({
  # find initial and final values for each metric for each person
  # one person at a time and then row bind the dataframes
  # there is a better way of doing this, but for now this is ok
  df_change_amit <- df %>%
    gather(key, value, -name, -date) %>%
    filter(name=="Amit") %>%
    filter(date==min(date) | date==max(date)) %>%
    spread(date, value)

  # subtract the first and last values for each metric to get absolute change
  # in each metric
  colnames(df_change_amit) <- c("name", "metric", "value_at_start", "value_at_end")
  df_change_amit <- df_change_amit %>%
    mutate(abs_change = abs(value_at_end-value_at_start)) %>%
    select(-value_at_end, -value_at_start)
  
  # repeat for person 2
  df_change_nidhi <- df %>%
    gather(key, value, -name, -date) %>%
    filter(name=="Nidhi") %>%
    filter(date==min(date) | date==max(date)) %>%
    spread(date, value)
  colnames(df_change_nidhi) <- c("name", "metric", "value_at_start", "value_at_end")
  
  # notice that we are setting the abs value to negative
  # this is because we want to display the chart such that one person's value
  # are to the left of zero and the other person to the right of zero
  # the x axis tick labels are all set to positive so that what the viewer
  # see is only a positive value indicating absolute change
  df_change_nidhi <- df_change_nidhi %>%
    mutate(abs_change = -abs(value_at_end-value_at_start)) %>%
    select(-value_at_end, -value_at_start)

  # bind the dataframes to get one dataframe with both person's data
  df_change <- bind_rows(df_change_amit, df_change_nidhi)

  # we want the metics in a particular display order
  df_change$metric <- factor(df_change$metric, levels = c("weight", 
                                                          "bmi", 
                                                          "total-lean-mass-percentage",
                                                          "body-fat-percentage",
                                                          "hydration-percentage"))
  df_change <- df_change %>%
    arrange(desc(metric))

  # X Axis Breaks and Labels 
  # the -15 to 15 range is by looking at the data
  brks <- seq(-15, 15, 1)
  lbls = paste0(as.character(c(seq(15, 0, -1), seq(1, 15, 1))), "")

  # Plot, see http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html for reference
  ggplot(df_change,
         aes(x = metric, y = abs_change, fill = name)) +   # Fill column
    geom_bar(stat = "identity", width = .6) +   # draw the bars
    scale_y_continuous(breaks = brks,   # Breaks
                       labels = lbls) + # Labels
    coord_flip() +  # Flip axes
    #labs(title="Every metric compared") +
    ylab("Absolute Change") +
    xlab("") + 
    theme_bw() + 
    theme(plot.title = element_text(hjust = .5), 
          axis.ticks = element_blank(), legend.title = element_blank()) + 
    scale_fill_tableau()  # Color palette
})

```